### 2025-10-20 YEAR2/MAT292_LectureNotes/main.tex
oct 20

\begin{definition}[Characteristic Equation]
    The characteristic equation of a second-order linear ODE with constant coefficients:
    $$y'' + p y' + q y = 0$$
    is given by the quadratic equation:
    \begin{equation}
    r^2 + pr + q = 0
    \end{equation}
    where \( r \) represents the roots of the characteristic polynomial. The nature of the roots (real and distinct, real and repeated, or complex conjugates) determines the form of the general solution to the ODE. This comes from the assuned solution of the form \( y = e^{rt} \).
\end{definition}

@@

\end{document}

>>>>

\end{example}
\subsection{Second Order Homogeneous ODE}
\begin{definition}[Characteristic Equation]
    The characteristic equation of a second-order linear ODE with constant coefficients:
    $$y'' + p y' + q y = 0$$
    is given by the quadratic equation:
    \begin{equation}
    r^2 + pr + q = 0
    \end{equation}
    where \( r \) represents the roots of the characteristic polynomial. The nature of the roots (real and distinct, real and repeated, or complex conjugates) determines the form of the general solution to the ODE. This comes from the assuned solution of the form \( y = e^{rt} \).
\end{definition}

@@

\subsection{Second Order Inhomogeneous ODE}
\begin{theorem}[General Solution of Inhomogeneous ODE]
    The general solution of a second-order inhomogeneous ODE:
    $$y'' + p y' + q y = \sum_i g_i(t)$$
    is given by:
    \begin{equation}
    y(t) = c_1 y_1(t) + c_2 y_2(t) + \sum_i y_{p_i}(t)
    \end{equation}
    where \(c_1, c_2\) are constants determined by initial conditions, \( y_h(t) = \{ y_1(t) , y_2(t) \} \) is the general solution of the corresponding homogeneous equation \( y'' + p y' + q y = 0 \), and \( y_{p_i}(t) \) are particular solutions corresponding to each \( g_i(t) \). Each \( y_{p_i}(t) \) satisfies:
    \begin{enumerate}
        \item $\text{Null}(L) = \{y_h\}$
        \item $L(y_{p_i}) = g_i(t)$
    \end{enumerate}
\end{theorem}
\begin{theorem}[Uniqueness of Particular Solutions]
    If \( y_{p_1}(t) \) and \( y_{p_2}(t) \) are two particular solutions of the inhomogeneous ODE:
    $$y'' + p y' + q y = g(t)$$
    then their difference \( y_{p_1}(t) - y_{p_2}(t) \) is a solution of the corresponding homogeneous equation:
    $$y'' + p y' + q y = 0$$

    And since the homongenous solution is unique up to a linear combination of basis solutions, we have:
    $$y_{p_1}(t) - y_{p_2}(t) = c_1 y_1(t) + c_2 y_2(t)$$
    where \( c_1 \) and \( c_2 \) are constants, and \( y_1(t) \) and \( y_2(t) \) are the fundamental solutions of the corresponding homogeneous equation.
\end{theorem}
\begin{proof}
    We have:
    \begin{align*}
        L(y_{p_1}) &= g(t) \\
        L(y_{p_2}) &= g(t)
    \end{align*}
    Subtracting these two equations gives:
    \begin{align*}
        L(y_{p_1}) - L(y_{p_2}) &= g(t) - g(t) \\
        L(y_{p_1} - y_{p_2}) &= 0
    \end{align*}
    Thus, \( y_{p_1}(t) - y_{p_2}(t) \) is indeed a solution of the corresponding homogeneous equation.
\end{proof}

\begin{definition}[Wronskian Derivation of $y_p$]
    If $y_1(t)$ and $y_2(t)$ are two linearly independent solutions of the corresponding homogeneous equation (a.k.a. the fundamental set of homogeneous solutions), then we can assume a particular solution of the form:
    $$y_p(t) = -y_1(t) \int \frac{y_2(t) g(t)}{W(t)} dt + y_2(t) \int \frac{y_1(t) g(t)}{W(t)} dt$$
    where \( W(t) = y_1(t)y_2'(t) - y_2(t)y_1'(t) \) is the Wronskian determinant of \( y_1 \) and \( y_2 \).
\end{definition}

\begin{example}
    Consider the following second-order inhomogeneous ODE:
    $$y'' - 3y'- 4y = 3e^{2t} + \sin(t)$$
    The corresponding homogeneous equation is:
    $$y'' - 3y' - 4y = 0$$
    The characteristic equation is:
    $$r^2 - 3r - 4 = 0$$
    Solving this gives the roots \( r_1 = 4 \) and \( r_2 = -1 \). Thus, the general solution of the homogeneous equation is:
    $$y_h(t) = C_1 e^{4t} + C_2 e^{-t}$$
    where \( C_1 \) and \( C_2 \) are constants determined by initial conditions.

    Next, we find particular solutions for each term on the right-hand side of the inhomogeneous equation. For the term \( 3e^{2t} \), we can assume a particular solution of the form $y_{p_1}(t) = A e^{2t}$. Substituting this into the left-hand side of the ODE gives:
    \begin{align*}
        y_{p_1}'' - 3y_{p_1}' - 4y_{p_1} &= 4A e^{2t} - 6A e^{2t} - 4A e^{2t} \\
        &= -6A e^{2t}
    \end{align*}
    Setting this equal to \( 3e^{2t} \) gives:
    $$-6A e^{2t} = 3e^{2t} \implies A = -\frac{1}{2}$$
    Thus, we have:
    $$y_{p_1}(t) = -\frac{1}{2} e^{2t}$$
    For the term \( \sin(t) \), we can assume a particular solution of the form \( y_{p_2}(t) = B \cos(t) + C \sin(t) \). Substituting this into the left-hand side of the ODE gives:
    \begin{align*}
        y_{p_2}'' - 3y_{p_2}' - 4y_{p_2} &= -B \cos(t) - C \sin(t) + 3B \sin(t) - 3C \cos(t) - 4B \cos(t) - 4C \sin(t) \\
        &= (-5B - 3C) \cos(t) + (3B - 5C) \sin(t)
    \end{align*}
    Setting this equal to \( \sin(t) \) gives the system of equations:
    \begin{align*}
        -5B - 3C &= 0 \\
        3B - 5C &= 1
    \end{align*}
    Solving this system gives \( B = -\frac{3}{34} \) and \( C = \frac{5}{34} \). Thus, we have:
    $$y_{p_2}(t) = -\frac{3}{34} \cos(t) + \frac{5}{34} \sin(t)$$
    Finally, the general solution of the inhomogeneous ODE is:
    $$y(t) = C_1 e^{4t} + C_2 e^{-t} - \frac{1}{2} e^{2t} - \frac{3}{34} \cos(t) + \frac{5}{34} \sin(t)$$
\end{example}

\paragraph{Operator Theory} We call the differential operator \( L = D^2 + pD + qI \) where \( D \) is the differentiation operator and \( I \) is the identity operator. Then, we would call $L: C^2(\mathbb{R}) \to C(\mathbb{R})$ a transformation from the space of twice differentiable functions to the space of continuous functions. We can consider the factorization of this operator:
$$L = (D - \lambda_1 I)(D - \lambda_2 I)$$
where \( \lambda_1 \) and \( \lambda_2 \) are the roots of the characteristic equation.

Then, we can consider the eigenvalues and eigenfunctions of this operator:
\begin{definition}[Eigenvalues and Eigenfunctions of Differential Operators]
    An eigenvalue \( \lambda \) of a differential operator \( L \) is a scalar such that there exists a non-trivial function \( y(t) \) (called an eigenfunction) satisfying the equation:
    \begin{equation}
    L[y(t)] = \lambda y(t)
    \end{equation}
    In the context of second-order linear ODEs, if we consider the operator \( L = D^2 + pD + qI \), then the eigenvalue problem becomes:
    \begin{equation}
    y'' + py' + qy = \lambda y
    \end{equation}
    The solutions to this equation can provide insights into the behavior of the system described by the differential equation, including stability and oscillatory properties.
\end{definition}

\begin{example}[Eigenfunctions of $D$]
    Consider the differential operator \( D \) defined by \( D[y(t)] = y'(t) \). We want to find the eigenvalues and eigenfunctions of this operator. The eigenvalue equation is given by:
    $$D[y(t)] = \lambda y(t)$$
    which translates to the first-order ODE, solving this gives:
    $$ f_\lambda = e^{\lambda t} $$
    where \( \lambda \) is any complex number. Thus, the eigenfunctions of the operator \( D \) exists in a dense, uncoutable spectrum.

    If we impose boundary conditions, for example, \( y(0) = 1 = y(2\pi) \), then we have:
    $$ e^{\lambda 2\pi} = 1 \implies \lambda = ik, k \in \mathbb{Z} $$
    Thus, the eigenvalues are discrete and countable, and the corresponding eigenfunctions are:
    $$ f_k(t) = e^{ikt}, k \in \mathbb{Z} $$
    which form a basis for the space of periodic functions with period \( 2\pi \).
\end{example}

\begin{definition}[Fourier Basis]
    The set of functions \( \{ e^{ikt} \mid k \in \mathbb{Z} \} \) forms an orthogonal basis for the space of square-integrable functions on the interval \( [0, 2\pi] \). This basis is known as the Fourier basis. Any function \( f(t) \) in this space can be expressed as a Fourier series:
    \begin{equation}
    f(t) = \sum_{k=-\infty}^{\infty} c_k e^{ikt}
    \end{equation}
    where the coefficients \( c_k \) are given by:
    \begin{equation}
    c_k = \frac{1}{2\pi} \int_0^{2\pi} f(t) e^{-ikt} dt
    \end{equation}
    The Fourier basis is widely used in various fields, including signal processing, heat transfer, and quantum mechanics, due to its ability to represent periodic functions and analyze frequency components.

\begin{definition}[Generalized Eigenfunction]
    A generalized eigenfunction of a differential operator \( L \) associated with an eigenvalue \( \lambda \) is a function \( y(t) \) that satisfies the equation:
    \begin{equation}
    (L - \lambda I)^m[y(t)] = 0
    \end{equation}
    for some positive integer \( m \). Here, \( I \) is the identity operator, and \( m \) is the order of the generalized eigenfunction. Generalized eigenfunctions arise in cases where the operator \( L \) has repeated eigenvalues, and they extend the concept of eigenfunctions to include functions that may not satisfy the standard eigenvalue equation but still provide valuable information about the structure of the solution space.

\begin{definition}[Spectrum]
    The spectrum of a differential operator \( L \) is the set of all eigenvalues \( \lambda \) for which there exists a non-trivial solution \( y(t) \) (eigenfunction):
    \begin{equation}
        \mathrm{spec}(L) = \{ \lambda \in \mathbb{C} \mid L[y(t)] = \lambda y(t) \text{ has a non-trivial solution } y(t) \}
    \end{equation}
\end{definition}

\begin{definition}[Variation of Parameters]
    Variation of parameters is a method used to find a particular solution to a non-homogeneous linear differential equation. For a second-order ODE of the form:
    $$y'' + p(t)y' + q(t)y = g(t)$$
    where \( y_1(t) \) and \( y_2(t) \) are two linearly independent solutions of the corresponding homogeneous equation, a particular solution \( y_p(t) \) can be expressed as:
    \begin{equation}
    y_p(t) = u_1(t)y_1(t) + u_2(t)y_2(t)
    \end{equation}
    where \( u_1(t) \) and \( u_2(t) \) are functions to be determined. These functions satisfy the system of equations:
    \begin{align*}
        u_1'(t)y_1(t) + u_2'(t)y_2(t) &= 0 \\
        u_1'(t)y_1'(t) + u_2'(t)y_2'(t) &= g(t)
    \end{align*}
    Solving this system allows us to find \( u_1(t) \) and \( u_2(t) \), and thus the particular solution \( y_p(t) \).

\end{definition}
\end{document}

### 2025-10-20 YEAR2/PHY293_LectureNotes/main.tex
oct 20

    To find the coefficients \( A_n \), we use:
    \begin{equation}
        A_n = \frac{2}{L} \int_0^L f(x) \sin\left(\frac{n\pi x}{L}\right) \, dx
    \end{equation}
\end{definition}
\end{document}

>>>>

    To find the coefficients \( A_n \), we use:
    \begin{equation}
        A_n = \frac{2}{L} \int_0^L f(x) \sin\left(\frac{n\pi x}{L}\right) \, dx
    \end{equation}
\end{definition}

\chapter{Modern Physics}
\section{Quantum Mechanics}
\begin{definition}[Quantization]
    Quantization is the constraint that energy and other physical properties can only take on discrete values (\textit{quanta}) rather than a continuous range.
\end{definition}
\begin{definition}[Probabilistic Framework]
    Predictions in quantum mechanics are inherently probabilistic. The exact outcome of a measurement cannot be determined, only the probability of different outcomes.
\end{definition}

\begin{definition}[Black Body]
    A black body is an idealized physical object that absorbs all incident electromagnetic radiation, regardless of frequency or angle of incidence. It also emits radiation in a characteristic spectrum that depends solely on its temperature.
\end{definition}
\begin{definition}[Classical View of Black Body Radiation]
    According to classical physics, Reyleigh-Jeans law predicts that the energy density of black body radiation increases indefinitely with frequency, leading to the "ultraviolet catastrophe." The classical formula is given by:
    $$    u(\nu, T) = \frac{8\pi \nu^2 k_B T}{c^3} $$
    where \( u(\nu, T) \) is the energy density per unit frequency, \( \nu \) is the frequency, \( k_B \) is the Boltzmann constant, \( T \) is the absolute temperature, and \( c \) is the speed of light.
\end{definition}
\begin{definition}[Quantum View of Black Body Radiation]
    Max Planck resolved the ultraviolet catastrophe by proposing that electromagnetic energy is quantized. He introduced the concept that energy is emitted or absorbed in discrete packets called \textbf{quanta}, with energy given by:
    \begin{equation}
        E_n = n h f \quad n = 0, 1, 2, \ldots
    \end{equation}
    where \( h \) is Planck's constant and \( f \) is the frequency of the radiation. This leads to Planck's law for black body radiation. The resulting law for the black body spectrum is:
    \begin{equation}
        u(\nu, T) = \frac{8\pi h \nu^3}{c^3} \frac{1}{e^{\frac{h\nu}{k_B T}} - 1}
    \end{equation}
\end{definition}

\begin{definition}[Photoelectric Effect]
    The photoelectric effect is the phenomenon where electrons are emitted from a material (usually a metal) when it is exposed to light of sufficient frequency. The key observations of the photoelectric effect are:
    \begin{itemize}
        \item Electrons are emitted only if the incident light has a frequency above a certain threshold frequency \( f_0 \).
        \item The kinetic energy of the emitted electrons depends on the frequency of the incident light, not its intensity.
        \item The number of emitted electrons is proportional to the intensity of the incident light, provided the frequency is above the threshold.
    \end{itemize}
    The energy of the incident photons is given by:
    $$
        E = h f
    $$
    where \( h \) is Planck's constant and \( f \) is the frequency of the light. The maximum kinetic energy \( K_{max} \) of the emitted electrons is given by:
    $$
        K_{max} = h f - \phi
    $$
    where \( \phi \) is the work function of the material, representing the minimum energy required to liberate an electron from the surface.

\begin{definition}[Campton Effect]
    The Compton effect is the increase in wavelength (or decrease in energy) of X-rays or gamma rays when they are scattered by electrons. This phenomenon provides evidence for the particle-like behavior of light. The change in wavelength \( \Delta \lambda \) of the scattered photon is given by the Compton formula:
    \begin{equation}
        \Delta \lambda = \lambda' - \lambda = \frac{h}{m_e c} (1 - \cos \theta)
    \end{equation}
    where:
    \begin{itemize}
        \item \( \lambda \) is the initial wavelength of the photon,
        \item \( \lambda' \) is the wavelength after scattering,
        \item \( h \) is Planck's constant,
        \item \( m_e \) is the rest mass of the electron,
        \item \( c \) is the speed of light,
        \item \( \theta \) is the angle at which the photon is scattered.
    \end{itemize}

\end{definition}
\end{document}
